### 2018-08-13 papers/1_dmatpol/submit/paper-submit.tex
Citation added and switch to reprint mode

%%
%%   Version 3.1 of 31 July 2018.
%%
\documentclass[aip,jcp,preprint,amsmath,amssymb,floatfix]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{txfonts}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

% hyphenation
\usepackage{hyphenat}

% tables
%\usepackage{dcolumn}
\usepackage{multirow}
%\usepackage{hhline}
\usepackage{booktabs}
%\usepackage{longtable}

% mathematics
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
%\usepackage{bm}
\usepackage{mathrsfs}
\usepackage{upgreek}
%\allowdisplaybreaks

% symbols
%\usepackage{gensymb}
%\usepackage{textcomp}

%---------------------------------------------------
% happy integral
\newcommand{\rint}[1]{\mbox{\Large $ \int\limits_{\mbox{\tiny  $#1$}}$}}
% SHORTCUTS
%\newcolumntype{,}{D{.}{,}{2}}
\newcommand{\citee}[1]{\ensuremath{\scriptsize^{\citenum{#1}}}}
\newcommand{\HRule}{\rule{\linewidth}{0.2mm}}
% Quantum notation
\newcommand{\bra}[1]{\ensuremath{\bigl\langle {#1} \bigl\lvert}}
\newcommand{\ket}[1]{\ensuremath{\bigr\rvert {#1} \bigr\rangle}}
\newcommand{\braket}[2]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigr\rangle}}
\newcommand{\tbraket}[3]{\ensuremath{\bigl\langle {#1} \bigl\lvert {#2} \bigl\lvert {#3} \bigr\rangle}}
% Math
\newcommand{\pd}{\ensuremath{\partial}}
\newcommand{\DR}{\ensuremath{{\rm d} {\bf r}}}
%\newcommand{\BM}[1]{\ensuremath{\mbox{\boldmath${#1}$}}}
\newcommand{\BM}[1]{\bm{#1}}
% Chemistry (formulas)
\newcommand{\ch}[2]{\ensuremath{\mathrm{#1}_{#2}}}
% Math 
\newcommand{\VEC}[1]{\ensuremath{\mathrm{\mathbf{#1}}}}
% vector nabla
\newcommand{\Nabla}{\ensuremath{ \BM{\nabla}}}
% derivative
\newcommand{\FDer}[3]{\ensuremath{
\bigg(
\frac{\partial #1}{\partial #2}
\bigg)_{#3}}}
% diagonal second derivative
\newcommand{\SDer}[3]{\ensuremath{
\biggl(
\frac{\partial^2 #1}{\partial #2^2}
\biggr)_{#3}}}
% off-diagonal second derivative
\newcommand{\SSDer}[4]{\ensuremath{
\biggl(
\frac{\partial^2 #1}{\partial #2 \partial #3}
\biggr)_{#4}}}
% derivatives without bound
% derivative
\newcommand{\fderiv}[2]{\ensuremath{
\frac{\partial #1}{\partial #2}}}
% diagonal second derivative
\newcommand{\sderiv}[2]{\ensuremath{
\frac{\partial^2 #1}{\partial #2^2}
}}
% off-diagonal second derivative
\newcommand{\sderivd}[3]{\ensuremath{
\frac{\partial^2 #1}{\partial #2 \partial #3}
}}
% derivatives for tables
\newcommand{\fderivm}[2]{\ensuremath{
{\partial #1}/{\partial #2}}}
% diagonal second derivative
\newcommand{\sderivm}[2]{\ensuremath{
{\partial^2 #1}/{\partial #2^2}
}}
% off-diagonal second derivative
\newcommand{\sderivdm}[3]{\ensuremath{
{\partial^2 #1}/{\partial #2 \partial #3}
}}
% ERIs and OEIs
\newcommand{\OEIc}[3]{\ensuremath{\left(#1 \lvert #2 \rvert #3 \right)}}
\newcommand{\ERIc}[4]{\ensuremath{\left(#1 #2 \vert #3 #4 \right)}}

% Partial density and potential
\newcommand{\PartPot}[4]{\ensuremath{\frac{#1 #2}{\lvert #3-#4 \rvert }}}

% trace operator
\DeclareMathOperator{\Tr}{Tr}

% EU logo
\newcommand*{\InlineImage}[1]{%
    \raisebox{-.02\baselineskip}{%
        \includegraphics[
        height=\baselineskip,
        width=\baselineskip,
        keepaspectratio,
        ]{#1}%
    }%
}

%\draft % marks overfull lines with a black rule on the right

\begin{document}

%\preprint{AIP/123-DMS}

\title{One-Particle Density Matrix Polarisation Susceptibility Tensors}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}
%\thanks{Bulak}
\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}kiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\begin{abstract}

%Considering the solvent effect on the correlated electronic wavefunctions, especially
%those that describe the electronically excited state, is a great challenge of Quantum Chemistry.

The electric field\hyp{}induced change in the one\hyp{}electron density
has been expressed as a series of the one\hyp{}particle density matrix 
susceptibilities (DMS) interacting with the spatial distribution of electric field.
The analytic approximate expressions are derived at the Hartree\hyp{}Fock theory, 
which serves as a basis for construction of generalised model 
that is designed for an arbitrary form of wavefunction and any type of one\hyp{}particle density matrix.
It is shown that it is possible to accurately predict 
the changes in the one\hyp{}electron ground\hyp{}state density of water molecule 
in spatially uniform electric field, as well as in spatially non\hyp{}uniform 
electric field distribution generated by point charges. When both linear 
and quadratic terms with respect to the electric field are accounted for,
the electric field\hyp{}induced polarisation energies, dipole moments 
and quadrupole moments are quantitatively described by the present theory 
in electric fields ranging from weak to very strong (0.001 -- 0.07 a.u.).
It is believed that the proposed model could open new routes in quantum chemistry 
for fast and efficient calculations of molecular properties in condensed phases.
\end{abstract}

\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

\section{\label{s:1}Introduction}

Molecular\hyp{}level exploration of condensed phase phenomena becomes increasingly unavoidable
as technology and medicine continue to develop. Unfortunately, the computational cost of accurate
electronic structure calculations grows extremely fast with the number of particles, greatly limiting
the use of the state\hyp{}of\hyp{}the\hyp{}art first\hyp{}principles theories when large and complex 
chemical systems such as solutions and biomolecules come into play.\cite{
Xu.Guidez.Bertoni.Gordon.JCP.2018,Tomasi.Mennucci.Cammi.ChemRev.2005}
Due to this reason, it has been of particular importance nowadays to develop bridging theoretical techniques 
that link extremely accurate but localised pieces of information with certain knowledge of an atomistic 
organisation and intermolecular interaction on relatively large spatial scales, typically ranging beyond 
few nanometers.

The development of hybrid methods such as quantum mechanical\hyp{}molecular mechanical (QM/MM) technique 
was a revolution in computational chemistry which allows insightful theoretical studies of quantum mechanical
effects in biological systems because costly quantum mechanical calculations can be combined with lower\hyp{}level 
and much less expensive classical dynamics.\cite{Warshel.Levitt.JMolBiol.1976,Senn.Thiel.Angew.2009}
Combination of accurate \emph{ab initio} methods with extensive experimental and semi\hyp{}empirical
refinings led to development of a wide variety of force fields,\cite{Demerdash.Yap.T-Head-Gordon.AnnuRevPhysChem.2014}
whereas sophisticated partitioning of Hamiltonian or electron density enabled numerous schemes for efficient fragmentation
within a QM level.\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
%Gao.Truhlar.Wang.Mazack.Loffler.Provorse.Rehak.ACR.2014}

However, even with the modern advances in computational chemistry, accurate description of phenomena due 
to complex heterogeneous molecular environments remains a challenge in the following areas: (i) describing 
the electronically excited states (electronic solvatochromism),\cite{Barbati.JACS.2014,
Szabla.Sponer.Jiri.Gora.JPCL.2015,Bednarska.Zalesny.Tian.Murugan.Agren.Bartkowiak.Molecules.2017,
Jedrzejewska.Grabarz.Bartkowiak.Osmialowski.SpectChimActA.2018} 
(ii) understanding the vibrational properties of molecules in ground and electronically excited states
(vibrational or vibronic solvatochromisms).\cite{Blasiak.Londergan.Webb.Cho.ACR.2017,
Xu.Blasiak.Cho.Layfield.Londergan.JPCL.2018,Lewis.Gruenke.Oliver.Ballottari.Bassi.Fleming.JPCL.2016} 
Interface between QM and MM levels is usually realised by electrostatic embedding and only rarely 
the Pauli exclusion principle and dispersion are additionally included.\cite{List.Olsen.Kongsted.PCCP.2016}
Unfortunately, it is quite difficult to efficiently treat the environment\hyp{}induced potential operator
due to complicated nature of intermolecular interactions that theoretically cannot be solely envisioned 
as a sum of separate contributions coming from electrostatic, dispersion and Pauli 
effects.\cite{Jeziorski.Moszynski.Szalewicz.ChemRev.1994} 
For example,
one of the most frequently used, efficient and reasonably accurate \emph{ab initio} fragmentation methods, 
the effective fragment potential (EFP) methods, are limited to the Hartree\hyp{}Fock 
(HF)\cite{Roothaan.RevModPhys.1951,Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013} 
or density functional theory (DFT) levels\cite{Hohenberg.Kohn.PhysRev.1964,Kohn.Sham.PhysRev.1965,Nguyen.Pachter.Day.JCP.2014}.
Unfortunately, while they can approximately describe the environment composed of molecules in their ground states,
they cannot be used for highly correlated wavefunctions such as electronically excited states.

Most of the efforts to theoretically treat extended molecular aggregates are rooted in energy\hyp{}based 
philosophy. However, in the past two decades there has been considerable interest in density\hyp{}based approaches
of a wider scope and aims.\cite{Yang.Lee.JCP.1995,
Mukamel.Berman.JCP.2003,
Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,
Elliott.Burke.Cohen.Wasserman.PhysRevA.2010,
Mandado.Hermida-Ramon.JCTC.2011,Sun.Chan.ACR.2016,Hedegard.Reiher.JCTC.2016} 
In particular, the importance of one\hyp{}electron densities in chemistry is well known and acknowledged 
in the community,\cite{Holas.March.PhysRevA.1991,Yang.PhysRevLett.1991,Yang.Lee.JCP.1995} 
though obtaining the accurate electron density distribution 
has been an unresolved issue for most of chemical systems in condensed phases. Given that fact, it would be 
incredibly insightful to establish a first\hyp{}principles connection between the one\hyp{}particle density 
and the perturbation in the environment through introduction of \emph{density matrix susceptibility} (DMS) tensors. 
The main postulate of the present work is that the associated change in the one\hyp{}particle density matrix, 
$\delta {\bf D}$, can be represented in a product form of the generalised perturbation susceptibility ${\bf B}_i$
and the environment\hyp{}induced perturbation on a molecule, $\delta {\bf P}_i$, evaluated at certain site $i$,
%
\begin{equation*} \label{e:initial-model}
  \delta {\bf D} \approx \sum_i^{\text{Sites}} {\bf B}_i
  \cdot  \delta {\bf P}_i \;,
\end{equation*}
%
where the `$\cdot$' symbol denotes the inner tensor (dot) product.
While $\delta {\bf P}_i$ contains critical information about the actual distortion of the electronic cloud,
perturbation susceptibility is assumed to be a property of unperturbed molecule and is to be precomputed only once. 
Thus, such DMS could be adapted into an efficient effective one\hyp{}electron fragment potential which in principle, 
is applicable to an arbitrary level of theory, as long as the one\hyp{}particle densities are available by means of
quantum chemistry methods. It also means that such a technique would dramatically reduce costs of quantum chemistry 
calculations because the task of performing full computation on a molecular aggregate would reduce to much simpler
matrix operations.

The simplest type of perturbation, yet the most relevant for chemistry, is the electrostatic perturbation.
It is worth mentioning that, at the moment, there is no simple \emph{ab initio} method to compute 
polarisation\hyp{}induced electron deformation densities without considering the entire molecular 
aggregate by using full quantum chemistry calculations.\cite{
Ma.Politzer.JCP.2004,
Horn.Head-Gordon.JCP.2015,
Mandado.Hermida-Ramon.JCTC.2011}
In this work, we explore the susceptibility of the molecule's electronic density
to polarise due to external electrostatic perturbation from first principles. 
The deformation density is expressed in terms
of the one\hyp{}particle density matrix polarisation susceptibility tensors 
that are properties of isolated molecules and can be
extracted from quantum chemistry calculations. The focus on the density (matrix) instead of the energy
(scalar) is beneficial since it provides the directional tensor description
of the polarisation that could be combined with other effects such as Pauli repulsion 
or dispersion,\cite{Mandado.Hermida-Ramon.JCTC.2011}
in contrast to the standard interaction energy approaches. We show that accurate polarisation 
of the one\hyp{}particle density
matrix is achievable and can be realised as long as the one\hyp{}particle density matrices
are available. It is believed that the proposed method
could be a basis for a construction of an 
alternative approach in fragment\hyp{}based
modelling of extended molecular aggregates that is not limited to the ground state chemistry.

We introduce the necessary theory and derive the \emph{ab initio} density matrix susceptibility model
at the HF approximation\cite{Roothaan.RevModPhys.1951} 
in Section~\ref{s:2}. Subsequently, this model serves as a basis for generalisation to the 
arbitrary one\hyp{}electron density with electronic correlation which is derived in Section~\ref{s:2}. 
The implementation details in a computer code as well as the validation procedures are detailed
in Section~\ref{s:3}. We then discuss the performance of the models in Section~\ref{s:4}.
Finally, in Section~\ref{s:5} we conclude the presentation with a few 
remarks for future development.

\section{\label{s:2}Theory}

The electronic energy of a closed\hyp{}shell molecule is given by the time\hyp{}independent Schr{\"o}dinger equation,
%
\begin{equation}
 E^{(0)} = \tbraket{\Psi^{(0)}}{\mathscr{H}^{(0)}}{\Psi^{(0)}} \;,
\end{equation}
%
where $E$ stands for the total electronic energy, $\Psi$ the electronic wavefunction and $\mathscr{H}$
the Hamiltonian operator. In the above, the superscript $(0)$
denotes the reference or unperturbed state that is assumed to be known.
Under the influence of an external electrostatic perturbation, the total electronic energy
becomes
%
\begin{equation}
 E[{\bf F}({\bf r})] = \tbraket{\Psi}{\mathscr{H}^{(0)} + \mathscr{V}({\bf F}({\bf r}))}{\Psi}
\end{equation}
%
in which ${\bf F}({\bf r})$ is a non\hyp{}uniform electric field distribution.
The electric field\hyp{}induced energy change is an unknown functional 
of the associated change in the one\hyp{}electron density,
%
\begin{equation}\label{e:rho}
 \rho({\bf r}) = 2\sum_{\alpha\beta} D_{\alpha\beta} \phi_\alpha({\bf r}) \phi^*_\beta({\bf r}) \;.
\end{equation}
%
In Eq.~\eqref{e:rho} $\phi_\alpha({\bf r})$ is a basis function, which can be for instance 
an atomic orbital (AO) or a molecular orbital (MO), and $D_{\alpha\beta}$ is the corresponding
one\hyp{}particle density matrix.
If perturbation\hyp{}independent basis is chosen, the electric field\hyp{}induced change in the one\hyp{}particle density
is given by
%
\begin{equation}\label{e:rho-1}
 \delta \rho({\bf r}) = 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \phi_\alpha({\bf r}) \phi^*_\beta({\bf r}) \;.
\end{equation}
%
The electric field\hyp{}induced change in the ground state's one\hyp{}particle density 
follows its energy minimum of the perturbed state. In addition, all the perturbed 
excited state one\hyp{}electron densities must be excited state solutions of the Schr{\"o}dinger equation
with the exact energy\hyp{}minimised ground state.

In the limit of a weak and spatially uniform electric field,
the polarisation\hyp{}induced density matrix change of any electronic state
or electronic transition is given by the following Taylor expansion:
%
\begin{equation}\label{e:dD-Taylor}
 \delta D_{\alpha\beta} = \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}}  \cdot {\bf F} 
   + \frac{1}{2} 
     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} : {\bf F} \otimes {\bf F}
   + \ldots 
\end{equation}
%
where the `$\otimes$' symbol denotes the outer tensor product.
In the above expression, the density matrix susceptibilities 
are simply equal to the derivatives of the density matrix
with respect to electric field (up to a multiplicative constant).
In a general case, i.e., when electric field is relatively strong 
and non\hyp{}uniform, single\hyp{}centred expansions, like the one 
from Eq.~\eqref{e:dD-Taylor},
%although valid for an arbitrary level of theory in uniform fields,
are not sufficient enough to accurately describe the local polarisation
of electron density distribution.

The development of a general first\hyp{}principles theory
that is capable of accurately capturing the electric field\hyp{}dependence 
of the one\hyp{}particle density becomes extremely difficult. This is not 
only because of the very complicated (and generally unknown) solution
to the electronic correlation problem, but also because a working model 
has to accurately predict all the $n^2$ elements of the density matrix
where $n$ is the size of basis set. Even relatively small errors
in a particular matrix element will accumulate when performing two\hyp{}
and four\hyp{}index contractions, and in turn, affect the evaluated properties 
with non\hyp{}negligible errors.

Due to all of these reasons, we adopt the following approach: 
we first develop a qualitative \emph{ab initio} distributed\hyp{}site 
model at the Hartree\hyp{}Fock level of theory. Further on, we generalise it 
to an arbitrary level and type of density matrix
using the functional form that was found by considering HF level as a guideline,
but treating the density matrix susceptibility tensors as certain functions 
of adjustable parameters.

\subsection{Ab Initio Density Matrix Susceptibilities: Hartree-Fock Level}

\subsubsection{Induced Dipole Moment}

$\delta {\bf D}$ from Eq.~\eqref{e:rho-1} should 
be directly related to the
polarisation\hyp{}induced multipole moments because they are
a measure of the distortion of the electron density distribution due to the external electric
field. 
Therefore, we start our analysis by considering the induced 
ground\hyp{}state electronic dipole moment of a closed\hyp{}shell molecule, 
%
\begin{equation} \label{e:dmu-exact}
 \delta {\BM{\upmu}}({\bf r}_Q) = 
     2 \Tr{ 
         \left[ 
              \delta {\bf D} \cdot {\mathbb{M}({\bf r}_Q)}
         \right] } \;,
\end{equation}
%
where ${\mathbb{M}({\bf r}_Q)}$ is atomic orbital (AO)\hyp{}representation
of the dipole moment operator defined with respect to origin at ${\bf r}_Q$,
%
\begin{equation}\label{e:m}
 \left[ {\mathbb{M}({\bf r}_Q)} \right]_{\alpha\beta} = -\tbraket{\alpha}{{\bf r} - {\bf r}_Q}{\beta} \;.
\end{equation}
%
(In this work we shall use the double struck font for matrices of vectors, so that
in this case ${\mathbb{M}}$ can be considered as a $n \times n$ matrix of vectors of length 3).
Note that if the total charge is conserved, the induced moments are independent of the origin.

In the Hartree\hyp{}Fock theory,\cite{Roothaan.RevModPhys.1951} ${\bf D}$ is approximated as
%
\begin{equation}
 {\bf D} = {\bf C} \cdot {\bf C}^\dagger \;,
\end{equation}
%
where ${\bf C}$ describes the occupied molecular orbitals (MO's) as linear combinations
of AO's, referred here to as the LCAO-MO matrix.
${\bf D}$ is an idempotent matrix and if expressed in orthogonal basis the following holds
%
\begin{equation}
 {\bf D}^2 = {\bf D} \;.
\end{equation}
%
The change in the molecular orbitals can be parameterised as suggested by McWeeny\cite{McWeeny.RevModPhys.1960}
by defining the idempotency\hyp{}preserving variation
%
\begin{equation}
 \delta {\bf C} = {\BM\Delta} \cdot {\bf C}^{(0)} \;,
\end{equation}
%
where $\BM\Delta$ is a square non\hyp{}singular and non\hyp{}symmetric matrix.
McWeeny showed that
%
\begin{equation} \label{e:dmat-change.exact}
 \delta {\bf D} = -{\bf D}^{(0)} + \left[ {\bf D}^{(0)} + {\bf v} \right] \cdot
                                   \left[ {\bf 1} + {\bf v}^\dagger \cdot {\bf v} \right]^{-1} \cdot
                                   \left[ {\bf D}^{(0)} + {\bf v}^\dagger \right] \;,
\end{equation}
%
where the auxiliary vector ${\bf v}$ is defined as
%
\begin{equation}
 {\bf v} \equiv \left[ {\bf 1} - {\bf D}^{(0)} \right] \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  \;.
\end{equation}
%
Expanding Eq.~\eqref{e:dmat-change.exact} in a Taylor series around ${\bf v}={\bf 0}$ and
truncating it at linear terms with respect to ${\BM\Delta}$ gives
%
\begin{subequations} 
 \begin{align}
 \delta {\bf D} &\cong \left[ {\bf 1} - {\bf D}^{(0)} \right] \cdot {\BM\Delta} \cdot {\bf D}^{(0)} + 
                        {\bf D}^{(0)} \cdot {\BM\Delta}^\dagger \cdot \left[ {\bf 1} - {\bf D}^{(0)} \right]  
 \label{e:dD-1} \\  &= 
  \delta {\bf C} \cdot {\bf C}^{(0)\dagger} + {\bf C}^{(0)} \cdot \delta {\bf C}^\dagger \nonumber \\
           &\qquad - {\bf D}^{(0)} \cdot \delta {\bf C} \cdot {\bf C}^{{(0)}\dagger} 
                   - {\bf C}^{(0)} \cdot \delta {\bf C}^\dagger \cdot {\bf D}^{(0)}  \;.
 \label{e:dD-2}
 \end{align}
\end{subequations}
%
The latter equality is true in a special case when the basis set functions are orthogonal, i.e., 
${\bf C}^{(0)\dagger} \cdot {\bf C}^{(0)}={\bf 1}$.
%Such a condition is necessary in order to meet the charge conservation
%requirement (see Appendix~\ref{a:orig-dep} for a detailed discussion).
%Therefore orthogonality of AO's is assumed from that point, unless stated otherwise.
%Also, from this point onwards we assume that the molecular orbitals 
%and the transformation matrix $\BM \Delta$ are real.
By substituting the McWeeny's result from Eq.~\eqref{e:dD-1}
into Eq.~\eqref{e:dmu-exact} one can show that
%
%\begin{multline} \label{e:dmu-4-exact.linear-approximation}
% \frac{1}{2} 
% \delta {\BM{\upmu}}({\bf r}_Q)
%  \cong
%   \Tr{ 
%    \left[ 
%         {\mathbb{M}} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}  
%    \right] }
%%
%  +\Tr{ 
%    \left[ 
%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger}
%    \right] } \\
%%
%  -\Tr{ 
%    \left[ 
%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta} \cdot {\bf D}^{(0)}
%    \right] }
%%
%  -\Tr{ 
%    \left[ 
%         {\mathbb{M}} \cdot {\bf D}^{(0)} \cdot {\BM\Delta}^{\dagger} \cdot {\bf D}^{(0)}
%    \right] } \;,
%\end{multline}
%%
%which can be recast to
%
\begin{equation}\label{e:dmu-l-vector}
 \frac{1}{4} 
 \delta {\BM{\upmu}}({\bf r}_Q)
  \cong
  \Re \left\{ 
 \Tr{ 
    \left[ 
         \delta {\bf C} \cdot \mathbb{L}({\bf r}_Q)
   \right] }  
  \right\} \;.
\end{equation}
%
In the above expression, the `$\Re$' symbol denotes the real part,
and the intermediate quantity
%
\begin{equation} \label{e:L-projector}
 \mathbb{L}({\bf r}_Q) \equiv {\bf C}^{(0)\dagger} \cdot \mathbb{M}({\bf r}_Q) \cdot \left( {\bf 1} - {\bf D}^{(0)} \right) \;.
\end{equation}
%
Note that $\mathbb{L}({\bf r}_Q)$ contains the projector onto
the unoccupied MO space at the absence of the external electric field. It means that the
perturbed solution is constructed from the unperturbed virtual orbitals as well.
%Note also the similarity between Eq.~\eqref{e:L-projector} and Eq.~\eqref{e:dmu-exact}.

For the sake of simplicity, 
the orthogonality of AO's is assumed in this work, unless stated otherwise.
It is proven in the Appendix~\ref{a:orig-dep} that under such circumstances the
induced dipole moment obtained in Eq.~\eqref{e:L-projector} is independent of the origin. 
This guarantees that the theory fulfils the total charge conservation requirement and,
therefore, in later course of the paper, the dipole origin will be omitted.

Although Eq.~\eqref{e:dmu-l-vector}
is exact up to first order in ${\BM\Delta}$, it cannot be inverted at this moment to obtain $\delta {\bf C}$.
If the orbitals are real, the analytic solution exists
provided that the electric field is uniform. In a more general case, i.e., when the electric field
is non\hyp{}uniform, the analytic solution cannot be found without further approximations.
Let the induced dipole moment be partitioned into separate contributions
from the occupied localised MO's (LMO's), i.e.,
%
\begin{equation} \label{e:dmu-part}
 \delta {\BM{\upmu}} = \sum_i^{\rm occ} \delta {\BM{\upmu}}_i \;,
\end{equation}
%
where the induced dipole moment associated with the $i$th LMO
is assumed to be linearly dependent on the local electric field, i.e.,
%
\begin{equation} \label{e:dmu-dpol}
 \delta {\BM{\upmu}}_i \approx {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) \;.
\end{equation}
%
In Eq.~\eqref{e:dmu-dpol}, ${\BM{\alpha}}_i$ is the distributed dipole\hyp{}dipole polarisability tensor
associated with the $i$th LMO whereas ${\bf F}({\bf r}_i)$ 
is the electric field evaluated at the $i$th LMO charge centroid,
${\bf r}_i = \tbraket{\phi_i}{\hat{\bf r}}{\phi_i}$.
By inserting the above result into Eq.~\eqref{e:dmu-l-vector},
recasting the right hand side in a matrix form
and assuming real orbitals,
one can separate the contributions coming from the occupied MO's as
%
\begin{equation} \label{e:dmu-l-vector-mo-transform-X}
 \frac{1}{4} {\BM{\alpha}}_i \cdot {\bf F}({\bf r}_i) 
   =
   \delta {\bf c}_i^{\rm T} \cdot {\bf L}_i \;.
\end{equation}
%
Note that now ${\bf L}_i \equiv \left[ \mathbb{L} \right]_i$ 
is a real rectangular matrix of dimension $n \times 3$ and $\delta{\bf c}_i$ is the $i$th column of the
change in the LCAO-MO matrix due to the polarisation process.
The solution for $\delta {\bf C}$ is
%
\begin{equation} \label{e:mu-ind.matrix.linear.solution}
  \delta {\bf c}_i^{\rm T} = \frac{1}{4}
                    {\bf F}^{\rm T}({\bf r}_i) \cdot {\BM{\alpha}}_i^{\rm T} \cdot
                    \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \;,
%                   {\BM{\alpha}}_i : {\bf F}({\bf r}_i)  \otimes
%                   \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \;,
\end{equation}
%
where $\left[ {\bf L}_i  \right]^{-1}_{\rm Left}$ is a left inverse
of ${\bf L}_i$ matrix,
%
\begin{equation} 
      \left[ {\bf L}_i  \right]^{-1}_{\rm Left}   \equiv
       \left[ {\bf L}_i^{\rm T} \cdot {\bf L}_i \right]^{-1} \cdot {\bf L}_i^{\rm T}   
\end{equation}
%
with the dimension of $3\times n$. Note that the right inverse of ${\bf L}_i$
does not exist because ${\bf L}_i \cdot {\bf L}_i^{\rm T}$ is singular due to the fact that usually
$n > 3$.

\subsubsection{Density Matrix Change}

Now we can obtain the final expression for the density matrix change 
due to external electric field that reproduces the 
induced dipole moment of the molecule.
The change in the LCAO\hyp{}MO coefficient is parameterised as
%
\begin{equation} \label{e:dC-dmatpol}
 \delta C_{\alpha i} = {\bf b}_{\alpha}^{(i;1)} \cdot {\bf F}({\bf r}_i)  \;,
\end{equation}
%
where the susceptibility tensor is defined as
%
\begin{equation} \label{e:susceptibility-b}
  b_{\alpha;w}^{(i;1)}= \frac{1}{4} \sum_u^{x, y, z} \left[ {\BM{\alpha}}_i \right]_{uw}
   \left[ \left[ {\bf L}_i  \right]^{-1}_{\rm Left} \right]_{u;\alpha}  
\end{equation}
%
for $w=x,y,z$. 
The superscript `$(1)$' underlines the fact that the susceptibility introduced here
is first\hyp{}order with respect to the electric field.
The result from Eq.~\eqref{e:dC-dmatpol} can be now inserted into 
the expression for $\delta {\bf D}$ from Eq.~\eqref{e:dD-2} to finally give
%
\begin{equation}\label{e:final-model.HF}
 \delta D_{\alpha\beta} \approx \sum_i^{\rm occ} {\bf B}^{(i;1)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  \;,
\end{equation}
%
where the \emph{density matrix polarisation susceptibility tensor} is
%
%\begin{multline}  \label{e:susceptibility-B} % PREPRINT
\begin{equation}  \label{e:susceptibility-B}
 {\bf B}^{(i;1)}_{\alpha\beta} = 
                               C_{\alpha i} {\bf b}_{\beta}^{(i;1)} + C_{\beta i} {\bf b}_\alpha^{(i;1)} %\\ PREPRINT
                                - \sum_\gamma 
                                 \left( 
               D_{\alpha\gamma} C_{\beta i} + D_{\beta\gamma} C_{\alpha i}
                                 \right)
                                           {\bf b}_{\gamma}^{(i;1)}
 \;.
\end{equation}
%\end{multline} % PREPRINT
%
Hence, the three dimensional vector ${\bf B}^{(i;1)}_{\alpha\beta}$ 
describes the local linear polarisation affinity 
of the density matrix element 
associated with the $i$th LMO centroid as well as the $\alpha$th and $\beta$th orthogonal AO's.
Transforming these vectors to a non\hyp{}orthogonal AO basis is straightforward.
It is emphasised here that all the elements of ${\bf B}_{\alpha\beta}^{(i;1)}$
are unique properties of electronically unperturbed molecule. 
However, they are valid only under the Hartree\hyp{}Fock
approximation, with a guarantee that only the overall charge (exactly) and the dipole moment (approximately,
limited by the flavour of the distributed polarisability approximation), 
but not higher multipole moments and polarisation energy, are correctly described.

Nevertheless, one very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
the density matrix increment is dependent only on the susceptibility element associated with the same
pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
vector susceptibility, different for each density matrix element. This feature will be used to
generalise the model to an arbitrary level of theory and order of electric field response in the next section.

\subsection{Generalised Density Matrix Susceptibility Model}

\subsubsection{Functional form of the model}

Eq.~\eqref{e:final-model.HF}, which is the final result of the previous section, 
can be used as a guideline to construct a general model of density matrix polarisation.
Quadratic (and higher\hyp{}order) terms with respect to the electric field
can be added to the formulation, which leads to
%
%\begin{multline}\label{e:final-model.General} % PREPRINT
\begin{equation}\label{e:final-model.General}
 \delta D_{\alpha\beta} = \sum_{i }^M \Big\{
                                      {\bf B}^{(i ; 1)}_{\alpha\beta} \cdot {\bf F}({\bf r}_i)  %\\ PREPRINT
%                       + \sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)
                        +             {\bf B}^{(i ; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) 
                        + \ldots \Big\} \;.
\end{equation}
%\end{multline} % PREPRINT
%
In the above equation, the density matrix susceptibilities
${\bf B}^{(i; R)}_{\alpha\beta}$, distributed over $M$ sites,
denote the $R$-th order response with respect to the electric field.
Note that
${\bf B}^{(i; 1)}_{\alpha\beta}$ is a vector of length $3$ whereas ${\bf B}^{(i; 2)}_{\alpha\beta}$
is a $3\times 3$ matrix. Also, the distributed expansion centres
(labelled as $i$) are generalised to any point in the molecule.\footnote{More general
contribution due to square of electric field could 
be $\sum_{ij}^M {\bf B}^{(ij; 2)}_{\alpha\beta} : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_j)$. 
However, for simplicity, we neglect the contributions from different sites $i$ and $j$
in the present discussion and leave it for future development.
} Note that Eq.~\eqref{e:final-model.General} can refer to any type of one\hyp{}particle density matrix
such as ground\hyp{}state, excited\hyp{}state or transition density matrix.
However, the exact mathematical form of the generalised density matrix susceptibilities 
from Eq.~\eqref{e:final-model.General} is unknown.

\subsubsection{Determining the generalised susceptibility tensors by linear regression}

Let $\left\{ {\bf F}^{(1)}({\bf r}), {\bf F}^{(2)}({\bf r}), \ldots, {\bf F}^{(N)}({\bf r}), \ldots \right\}$ 
be a set of $N_{\rm max}$ distinct and randomly sampled 
spatial distributions of electric field. It is assumed that
the exact difference one\hyp{}particle density matrices (with respect to the unperturbed state)
defined as
%
\begin{equation}
 \delta \overline{\bf D}^{(N)} \equiv \overline{\bf D}^{(N)} - \overline{\bf D}^{(0)}
\end{equation}
%
are known for each sample (overline symbolises the exact estimate).
Now, for each pair of the AO indices we construct the following 
parameterisation of Eq.~\eqref{e:final-model.General}:
%
%\begin{multline}\label{e:final-model.General.Parameters} % PREPRINT
\begin{equation}\label{e:final-model.General.Parameters}
 \delta D^{(N)} = \sum_{i }^M \Big\{ 
                              \sum_u^{x,y,z} s^{[1]}_{iu} F_{iu}^{(N)} %\\ % PREPRINT
%               + \sum_{ij}^M \sum_{uw}^{x,y,z} \left( s^{[2]}_{iu} + s^{[2]}_{jw} \right) F_{iu}^{(N)} F_{jw}^{(N)}
                +             \sum_{u}^{x,y,z} \sum_{w<u} r_{uw} s^{[2]}_{iuw} F_{iu}^{(N)} F_{iw}^{(N)} 
                        + \ldots \Big\}
\end{equation}
%\end{multline} % PREPRINT
%
(the Greek subscripts were omitted here for notational simplicity).
In the above equation, $B_u^{(i;1)} = s^{[1]}_{iu}$ and $B_{uw}^{(i;2)} = r_{uw} s^{[2]}_{iuw}$,
where $r_{uw}$ is the symmetry factor equal to 1 for diagonal elements and 2 for off\hyp{}diagonal
elements of $B_{uw}^{(i;2)}$.
The multiple parameter blocks
(${\bf s}^{[1]}$, ${\bf s}^{[2]}$ and so on)
appear in the first power, allowing for linear least\hyp{}squares regression.
The square bracket superscripts denote the block of the parameter space.

To determine the optimum set,
$
 {\bf s} = 
\begin{pmatrix}
{\bf s}^{[1]} &
{\bf s}^{[2]} & \cdots
\end{pmatrix}^{\rm T}
$, we define a loss function $Z$ 
that is subject to the least\hyp{}squares minimisation,
%
\begin{equation}\label{e:Z}
 Z({\bf s}) = \sum_N^{N_{\rm max}} \left( \delta D^{(N)} - \delta \overline{D}^{(N)} \right)^2 \;.
\end{equation}
%
The Hessian of $Z$ computed with respect to the parameters 
is parameter\hyp{}independent (constant) and generally non\hyp{}singular
as long as the electric fields on all distributed sites are different.
Therefore, the exact solution for the optimal parameters is given by the Newton equation
%
\begin{equation}\label{e:Newton}
 {\bf s} = -{\bf H}^{-1} \cdot {\bf g} \;,
\end{equation}
%
where ${\bf g}$ and ${\bf H}$ are the gradient vector and the Hessian matrix, respectively.
The explicit forms of the gradient and Hessian up to second\hyp{}order 
are given in the Appendix~\ref{a:blocks}.
Note that in this case the dimensions of parameter space for the block 1 and 2 are
equal to $3M$ and $6M$, respectively.

\section{\label{s:3}Calculation Details}

The theory that is necessary to validate the density matrix susceptibility models from Section~\ref{s:2}
was implemented in our in\hyp{}house plugin to {\sc Psi4} quantum chemistry program.\cite{Psi4.JCTC.2017}
Water molecule was selected as a test system which was subjected to 
electrostatic perturbations due to random uniform electric fields and random collections of point charges
generating non\hyp{}uniform electric fields.
Components of spatially uniform electric fields were determined 
from the uniform distribution in the interval 
of $\vert 0.0404 \vert$ a.u. generating average field of 0.035 a.u. in an ensemble of 100 samples.
Cartesian coordinates of point charges were drawn from the uniform distribution 
of a set of forty three\hyp{}dimensional Euclidean points within a sphere of radius 6.0 Bohrs
with an origin that is located in the centre of mass of water molecule.
In selection of random points, space inside the union of spheres surrounding
oxygen and hydrogen atoms with radii of 4.5 and 4.0 Bohrs, respectively, was
kept free from any point charges due to potential problems with the convergence of the 
self\hyp{}consistent field (SCF) calculations.
Electric charge values were selected from the one\hyp{}dimensional uniform distribution
in the interval of $\vert q_{\rm max} \vert$, where $q_{\rm max}$
was tuned to obtain a desired ensemble of weak, moderate and strong electric field magnitudes
(with $q_{\rm max}=0.05$, $0.2$ and $0.4$ a.u., accordingly).
In this way, independent sets of 100 random samples of either uniform electric field or
point charge distributions were selected and exact solutions of HF\hyp{}SCF equations were 
performed for each validation procedure.
In the case of non\hyp{}uniform electric fields, the distribution centres 
of the \emph{ab initio} DMS model (Eq.~\eqref{e:final-model.HF}) were assumed 
to be the LMO centroids (resulting in the five\hyp{}centre model), 
while the atoms were used for the generalised DMS models (Eq.~\eqref{e:final-model.General}, 
resulting in the three\hyp{}centre models).
This is because certain LMO centroids are often located at the same point which would
introduce linear dependencies in the Hessian matrix from Eq.~\eqref{e:Newton.gH}
when using the generalised DMS scheme.

In order to validate all the models discussed in this contribution,
the electric field\hyp{}induced polarisation energy, 
dipole moment and quadrupole moment were analysed in detail.
The electric field\hyp{}induced polarisation energy at the HF level, $\delta E^{\rm Pol}$,
was computed from the following exact formula:
%
%\begin{multline}\label{e:dE} % PREPRINT
\begin{equation}\label{e:dE}
 \delta E^{\rm Pol}[{\bf F}] = 
                     \Tr{\left\{ {\bf D}^{(0)} \cdot
                                \left( 2\delta {\bf J} - \delta {\bf K}  \right) \right\}} %\\ % PREPRINT
                   + \Tr{\left\{ \delta {\bf D} \cdot
                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} + 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} \;.
\end{equation}
%\end{multline} % PREPRINT
%
In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
the one\hyp{}electron core Hamiltonian matrix of an unperturbed HF ground state,
${\bf V}$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
the polarisation of the electron density, respectively. 
In order to compare the performance of the usual 
distributed dipole\hyp{}dipole polarisability approach,\cite{Li.Netzloff.Gordon.JCP.2006}
the approximate interaction energy was also evaluated by using the expression:
%
\begin{equation}\label{e:dE-dpol}
 \delta E^{\rm Pol}[{\bf F}] \approx 
                   - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i) \;,
\end{equation}
%
in which ${\BM \alpha}_i$ is the distributed anisotropic dipole\hyp{}dipole polarisability tensor associated with the
$i$th LMO centroid (see Eq.~\eqref{e:dmu-dpol}).
The distributed polarisabilities were computed by utilising the coupled\hyp{}perturbed Hartree\hyp{}Fock
method.\cite{McWeeny.RevModPhys.1960,Dodds.McWeeney.Sadlej.MolPhys.1977} 
Orbital localisation was performed by using the Boys method.\cite{Boys.RevModPhys.1960}

The electric field\hyp{}induced dipole and primitive quadrupole moments were evaluated from the following expressions:
%
\begin{subequations}\label{e:dmult}
  \begin{align}
   \delta \upmu_{u}   &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{u}{\beta}  \;,\\
   \delta \Theta_{uw} &= 2\sum_{\alpha\beta} \delta D_{\alpha\beta} \tbraket{\alpha}{(u-u_{\rm CM})(w-w_{\rm CM})}{\beta} \;, 
  \end{align}
\end{subequations}
%
where the subscript `CM' denotes the centre of mass of water molecule.
When the distributed polarisabilities were used instead of the polarisation\hyp{}induced 
difference density matrices, primitive induced quadrupole moments were computed from
the induced dipole moments (given in Eq.~\eqref{e:dmu-dpol}) according to
%
\begin{equation}\label{e:dTheta-dpol}
  \delta {\BM \Theta} = \sum_i^{\rm occ} 
   \left\{ 
       \delta {\BM \upmu}_i \otimes \left( {\bf r}_i - {\bf r}_{\rm CM} \right) 
      + \left( {\bf r}_i - {\bf r}_{\rm CM} \right) \otimes \delta {\BM \upmu}_i
   \right\} \;.
\end{equation}
%
Primitive quadrupole moments were subsequently transformed to their traceless forms by
%
\begin{equation}
  \delta \widetilde{\Theta}_{uw} = \delta \Theta_{uw} - \frac{1}{3} \delta_{uw} \Tr{\BM\Theta} \;.
\end{equation}
%
Methodology for the calculation of the electric field due to point charges
can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}

To compute the interaction energy and the induced multipole moments, either exact or approximated forms
of the difference density matrix from Eq.~\eqref{e:dD-Taylor}, Eq.~\eqref{e:final-model.HF} or Eq.~\eqref{e:final-model.General},
were used for a particular electric field distribution. The statistical evaluation of
errors with respect to the benchmark (full HF calculations) were based on 
the following statistical quality descriptors:
%
\begin{enumerate}
 \item Root Mean Square Energy Error (RMSE)
   \begin{equation}\label{e:rmse}
     {\rm RMSE} = \sqrt{ \frac{1}{N_{\rm max}} \sum_N^{N_{\rm max}} \left( \delta E^{(N)} - \overline{\delta E^{(N)}}\right)^2 } \;;
   \end{equation}
 \item Root Mean Square Dipole Error (RMSD)
   \begin{equation}\label{e:rmsd}
     {\rm RMSD} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N^{N_{\rm max}} \left( \lvert \delta {\BM\upmu}^{(N)} \rvert 
                                             - \lvert \overline{\delta {\BM\upmu}^{(N)} \rvert }\right)^2 } \;;
   \end{equation}
 \item Root Mean Square Quadrupole Error (RMSQ)
   \begin{equation}\label{e:rmsq}
     {\rm RMSQ} = \sqrt{ \frac{1}{N_{\rm max}}\sum_N^{N_{\rm max}} \left( \lvert \delta \widetilde{\BM\Theta}^{(N)} \rvert 
                                             - \lvert \overline{\delta \widetilde{\BM\Theta}^{(N)} \rvert }\right)^2 } \;.
   \end{equation}
\end{enumerate}
%
The first describes the ability of the model to reproduce the exact energetics of the system, 
whereas the following two contain additional information on how well the deformation of the one\hyp{}electron
density is described. For Eqs.~\eqref{e:rmsd} and \eqref{e:rmsq} the absolute magnitudes
are defined as
%
\begin{subequations}\label{e:absmagn}
  \begin{align}
   \lvert {\BM\upmu} \rvert &= \sqrt{\mu_x^2+\mu_y^2+\mu_z^2} \;,\\
   \lvert {\widetilde{\BM\Theta}}\rvert &= 
          \sqrt{\widetilde{\Theta}_{zz}^2 + \frac{1}{3}\left(\widetilde{\Theta}_{xx}-\widetilde{\Theta}_{yy}\right)^2 
          + \frac{4}{3}\left( \widetilde{\Theta}_{xy}^2 + \widetilde{\Theta}_{xz}^2 + \widetilde{\Theta}_{yz}^2 \right)} \;. 
  \end{align}
\end{subequations}
%
Throughout the work, the Hartree\hyp{}Fock level\cite{Roothaan.RevModPhys.1951} was assumed 
and the 6-311++G** basis set\cite{Krishnan.Binkley.Seeger.Pople.JCP.1980}
was used. Before all the calculations, water molecule was energy\hyp{}optimised
with a maximum energy, force and displacement thresholds of 10$^{-6}$, 3$\times$$10^{-4}$ 
and 1.2$\times$$10^{-3}$ a.u., respectively.
For all the calculations of the density matrix polarisation susceptibility tensors
from Eqs.~\eqref{e:final-model.HF} and \eqref{e:susceptibility-B}, 
atomic basis was orthogonalised by using the L{\"o}wdin symmetric orthogonalisation scheme,\cite{Mayer.IJQC.2002}
and further back\hyp{}transformed to the non\hyp{}orthogonal AO basis.
To compute numerical first and second derivatives of the one\hyp{}particle density matrix
with respect to the electric field from Eq.~\eqref{e:dD-Taylor}, 
the following finite\hyp{}difference formulae were utilised:
%
\begin{subequations}\label{e:ff}
  \begin{align}
    D_u    &= \frac{D(F_u) - D(-F_u)}{2F_u} \;,\\
    D_{uu} &= \frac{D(F_u) + D(-F_u) - 2D^{(0)}}{F_u^2} \;,\\
    D_{uw} &= \frac{D(F_u,F_w) + D(-F_u,-F_w) + 2D^{(0)}}{2F_uF_w} \nonumber \\ 
           & \; -\frac{D(F_u) + D(-F_u) + D(F_w) + D(-F_w)}{2F_uF_w} \;,
  \end{align}
\end{subequations}
%
where 
the subscripts $u,w=x,y,z$ on the right\hyp{}hand sides denote the derivatives with respect 
to the Cartesian component of the electric field, and the finite difference
steps $F_{u(w)}$ were assumed to be 0.0005~a.u. for all components.

\section{\label{s:4}Results and Discussion}

\subsection{\label{ss:41}Water molecule in spatially uniform electric field}

Polarisation under the uniform electric field is the simplest type of electrostatic perturbation
whereby it is possible to compute the exact DMS according to Eq.~\eqref{e:dD-Taylor} 
directly by numerical differentiation. Here we consider DMS up to second order, i.e.,
%
\begin{subequations}\label{e:exact-DMS}
 \begin{align}
  {\bf B}^{(1)}_{\alpha\beta} &= \frac{\partial D_{\alpha\beta}}{\partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B10} \\
  {\bf B}^{(2)}_{\alpha\beta} &= \frac{1}{2} 
     \frac{\partial^2 D_{\alpha\beta}}{\partial {\bf F} \otimes \partial {\bf F}}\Bigg|_{{\bf F}={\bf 0}} \;, \label{e:B20}
 \end{align}
\end{subequations}
%
and electric field magnitudes in a range from very weak ($<$0.001 a.u.) to very strong ($\sim$0.07 a.u.).
The \emph{ab initio} model from Eq.~\eqref{e:final-model.HF} and
the generalised model from Eq.~\eqref{e:final-model.General} are also discussed for the sake of comparison. 
The associated error distributions in prediction of polarisation energies 
are shown in scatter plot in Figure~\ref{f:fig-1}, whereas the
statistical performance descriptors of polarisation energies and induced multipole moments
are gathered in Table~\ref{t:results}.
%
\begin{figure}[t]
\includegraphics[width=0.5\textwidth]{fig-1.eps}
\caption{\label{f:fig-1} {\bf Performance of the DMS tensors of the ground\hyp{}state water molecule
in uniform electric field.} 
$R_{\rm max}$ denotes the maximum rank of the electric field
in the expansion from Eq.~\eqref{e:final-model.General}.
Electric field magnitudes in a range from 0.0 to 0.07 a.u. were tested in 100 samples with randomly selected 
directions of electric field.
} 
\end{figure}
%

All the tested models that are linear with respect to the electric field,
i.e., from Eq.~\eqref{e:final-model.HF} (black solid squares in Figure~\ref{f:fig-1}) 
as well as truncated at linear susceptibilities $\mathbb{B}^{(1)}$ 
from Eqs.~\eqref{e:dD-Taylor} (blue solid circles) and \eqref{e:final-model.General} (blue open circles),
incorrectly reproduce the polarisation energy.
Interestingly, the predicted values are in strong
correlation with the exact estimates and exhibit similar slope of $\sim$2.
Incorporation of the second\hyp{}order term
obtained either by numerical differentiation from Eq.~\eqref{e:B20} (red filled circles) 
or by using the generalised DMS model (red open circles)
dramatically reduces the errors
leading to quantitative accuracy in all tested field magnitudes 
with the best RMSE of 0.173 kcal/mol ($R^2=99.1\%$) for the generalised DMS model.
On the other hand, the
induced dipole and quadrupole moments are reproduced reasonably well by
using the first\hyp{}order DMS with $R^2$ coefficients being 
greater than 97\% in case of dipole moments and from 69\% up to 81\% in case of quadrupole moments.
This observation can be understood given the fact that the induced dipole
and quadrupole moment is approximately proportional to the external electric field, whereas the polarisation
energy exhibits quadratic dependence (c.f. Eq.~\eqref{e:dE-dpol}), hence the need of including at least
quadratic terms in the expansion from Eq.~\eqref{e:dD-Taylor} or Eq.~\eqref{e:final-model.General}.

In comparison to the dipole\hyp{}dipole polarisability approach, 
quadratic DMS models perform only slightly worse 
(for example, RMSE = 0.173 and 0.112 kcal/mol when the generalised quadratic DMS
model and the polarisability model are used, respectively).
However, quadratic DMS models better predict the induced multipole moments with
%work better
%when predicting the induced multipole moments with
$R^2$ systematically greater than 99\%. 
Certainly, taking into account the distributed quadrupole polarisabilities
would greatly reduce errors of the standard distributed polarisability approach.
Nevertheless, it is clear that the second\hyp{}order
DMS models are quantitative even at very strong uniform electric fields.
%On the other hand, DMS models better predict the induced dipole and quadrupole 
%moments even in weak fields.

%
\begin{table*}%[]
\caption[Performance of the DMS tensors of ground\hyp{}state water molecule
in uniform and non\hyp{}uniform electric field]{\label{t:results}
{\bf Performance of the DMS tensors of ground\hyp{}state water molecule
in uniform and non\hyp{}uniform electric field}  
\footnotemark[1]\footnotemark[2]
}
%\begin{ruledtabular}
\resizebox{\textwidth}{!}{%
\begin{tabular}{cddddddcddddcdd}
\hline\hline
%\toprule
\multicolumn{1}{l}{} & \multicolumn{6}{c}{\textbf{Ab Initio DMS}} &  & \multicolumn{4}{c}{\textbf{Generalised DMS}}
                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
\cline{2-8}
\cline{9-12}
%\cline{14-15}
\textbf{}         
& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$\footnotemark[3]}
& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$\footnotemark[4]}
& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$\footnotemark[4]}
&
& \multicolumn{2}{c}{$\BM{R_{\rm max}=1}$\footnotemark[5]} 
& \multicolumn{2}{c}{$\BM{R_{\rm max}=2}$\footnotemark[5]} 
&  
& \multicolumn{2}{c}{\textbf{Polarisability}}        \\ \hline
\multicolumn{15}{c}{} \\ 
\multicolumn{15}{c}{\textbf{Uniform Electric Field}} \\ 
\multicolumn{15}{c}{} \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}              \\ %\hline
   & 4.041 & (n.d) & 4.823 & (n.d.) & 0.231 & (98.40) && 5.0121 & (n.d.) & 0.173 & (99.10) && 0.112 & (99.63) \\ 
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}                    \\ %\hline
   & 30.46 & (97.91) & 30.46 & (97.91) & 18.02 & (99.27) && 24.83 & (98.61) & 7.13 & (99.89) && 30.46 & (97.91) \\ 
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}} \\ %\hline
   & 47.89 & (69.17) & 39.24 & (79.30) &  8.56 & (99.02) && 37.63 & (80.96) & 3.94 & (99.79) && 50.66 & (65.49) \\ \hline
\multicolumn{15}{c}{} \\ 
\multicolumn{15}{c}{\textbf{Non-Uniform Electric Field}\footnotemark[6]} \\ 
\multicolumn{15}{c}{} \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}   \\ 
W & 0.067 & (n.d.)  & - &  -  & - &  -  && 0.076 & (n.d.)  & 0.003 & (99.45) && 0.004 & (98.64) \\ 
M & 1.072 & (n.d.)  & - &  -  & - &  -  && 1.236 & (n.d.)  & 0.057 & (99.08) && 0.070 & (98.58) \\ 
S & 4.220 & (n.d.)  & - &  -  & - &  -  && 5.147 & (n.d.)  & 0.402 & (97.25) && 0.332 & (98.13) \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}  \\ 
W &  3.36 & (98.71) & - &  -  & - &  -  &&  1.03 & (99.88) &  0.94 & (99.90) &&  3.36 & (98.71) \\ 
M & 14.72 & (98.50) & - &  -  & - &  -  &&  7.72 & (99.59) &  4.55 & (99.86) && 14.72 & (98.50) \\ 
S & 44.80 & (96.92) & - &  -  & - &  -  && 33.06 & (98.33) & 15.96 & (99.61) && 44.80 & (96.92) \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}} \\ 
W &  6.22 & (58.72) & - &  -  & - &  -  &&  1.59 & (97.32) &  1.42 & (97.86) &&  5.45 & (68.29) \\ 
M & 28.22 & (53.90) & - &  -  & - &  -  && 10.45 & (93.68) &  5.68 & (98.13) && 23.94 & (66.83) \\ 
S & 78.07 & (40.53) & - &  -  & - &  -  && 43.33 & (81.68) & 14.40 & (97.98) && 64.33 & (59.62) \\ 
%\bottomrule
\hline\hline
\end{tabular}%
}
%\end{ruledtabular}
\footnotetext[1]{$R$-squared coefficients are shown in parentheses. `$n.d.$' stands for `not determined'.}
\footnotetext[2]{Ranks of the density matrix susceptibility models are denoted as $R_{\rm max}$ where $R_{\rm max}$ is the maximum power in the expansion from Eq.~\eqref{e:final-model.General}.}
\footnotetext[3]{\emph{Ab initio} HF model, Eq.~\eqref{e:final-model.HF}.}
\footnotetext[4]{\emph{Ab initio} finite\hyp{}difference model, Eq.~\eqref{e:dD-Taylor}.}
\footnotetext[5]{Linear\hyp{}regression generalised model, Eq.~\eqref{e:final-model.General}.}
\footnotetext[6]{`W', `M' and `S' denote `weak', `moderate' and `strong' electric fields that are generated by sets of point charges, respectively (c.f. Table~\ref{t:fields}).}
\end{table*}

\subsection{\label{ss:42}Water molecule in spatially non\hyp{}uniform electric field}

To study the capability of the distributed DMS models to capture 
density matrix polarisation in inhomogeneous electric fields, a water molecule
surrounded by point charges is analysed.
Three different types of point charge environment that generate
weak (0.003 -- 0.007 a.u.), 
moderate (0.01 -- 0.02 a.u.) and strong (0.02 -- 0.05 a.u.) 
electric fields around the water molecule were considered (Table~\ref{t:fields}),
%
\begin{table}[b]
\caption[Average electric fields in statistical sets of electrostatically perturbed ground states
of water molecule surrounded by point charges]
{{\bf Average electric fields in statistical sets of electrostatically perturbed ground states
of water molecule surrounded by point charges\footnotemark[1]}
}
\label{t:fields}
\begin{ruledtabular}
\begin{tabular}{ldcdcdcd}
\multirow{2}{*}{\textbf{Set\footnotemark[2]}} 
                                  & \multicolumn{7}{c}{\textbf{Average Electric Field (a.u.)}}      \\
                                  & \multicolumn{3}{c}{\textbf{Oxygen Atom}} & \textbf{} 
                                  & \multicolumn{3}{c}{\textbf{Hydrogen Atom}} \\
\cline{2-4}
\cline{6-8}
\textbf{W}                    & 0.0046     & $\pm$     & 0.0019     &           & 0.0046     & $\pm$     & 0.0020     \\
\textbf{M}                    & 0.0184     & $\pm$     & 0.0075     &           & 0.0186     & $\pm$     & 0.0078     \\
\textbf{S}                    & 0.0368     & $\pm$     & 0.0150     &           & 0.0372     & $\pm$     & 0.0156    
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Each set was composed of 100 samples differing in the configuration of 40 charges generated from a uniform distribution.}
\footnotetext[2]{`W', `M' and `S' denote `weak', `moderate' and `strong' electric fields, respectively.}
\end{table}
%
because similar ranges of electric field are experienced by water molecules
in liquid phase.\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013,Reischl.Kofinger.Dellago.MolPhys.2009}
Performance descriptors of the \emph{ab initio} and generalised 
models of the density matrix susceptibility, as well as 
the conventional dipole\hyp{}dipole distributed polarisability
approach, are shown in Table~\ref{t:results}, whereas the error distributions
in strong electric fields are additionally plotted in Figure~\ref{f:fig-2}A--I.

Similarly as in uniform electric field, 
the \emph{ab initio} DMS model from Eq.~\eqref{e:final-model.HF} 
as well as the generalised linear DMS model, are insufficient to correctly describe the polarisation energy 
in non\hyp{}uniform electric field, although the correlation with benchmark is still very clear with a slope of $\sim$2
(Figure~\ref{f:fig-2}A). Again, induced dipole moments are reproduced very well also
in strong electric fields (RMSD of 44.8~mD and lesser, $R^2>96\%$; Figure~\ref{f:fig-2}D and \ref{f:fig-2}E).
Induced quadrupole moments are well described only by the generalised DMS model.
For example, in moderate electric fields RMSQ is 10.4$\times 10^{-3}$D\AA{ }with $R^2=93.7\%$
whereas the \emph{ab initio} DMS model yields RMSQ of 28.2$\times 10^{-3}$D\AA{ }($R^2=53.9\%$).

\begin{figure*}[h]
\includegraphics[width=\textwidth]{fig-2.eps}
\caption{\label{f:fig-2} {\bf Performance of the distributed DMS tensors of ground\hyp{}state water molecule in strong
non\hyp{}uniform electric field.}
In this Figure, `Ab Initio DMS' (panels A, D and G) 
refers to the model from Eq.~\eqref{e:final-model.HF} and
`Generalised DMS' (panels B, E and H) 
refers to the generalised model with $R_{\rm max}=2$ 
where $R_{\rm max}$ is the maximum power
in the expansion from Eq.~\eqref{e:final-model.General}.
Comparison with the distributed dipole\hyp{}dipole polarisability model %(Eq.~\eqref{e:dE-dpol})
is also shown in this Figure in panels C, F and I. 
The ensemble of electric fields corresponds to the `S' set
from Tables~\ref{t:results} and \ref{t:fields}.}
\end{figure*}

The most sophisticated model studied in this work, the generalised quadratic DMS model, yields excellent predictions 
of the polarisation energies and induced multipole moments in all strengths of the non\hyp{}uniform electric field tested.
The errors in predicting polarisation energy are reduced by over an order of magnitude when compared with
the linear DMS models. Improvement of accuracy can also be observed in the case of induced multipole moment predictions
but the effect is less significant.
%(improvement roughly by a factor of 2).
What is particularly interesting is that in weak and moderate electric fields
the generalised quadratic DMS model performs better than the distributed dipole\hyp{}dipole
polarisability model, whereas only in strong electric fields the situation is reversed
(RMSE of 0.402, $R^2=97.2\%$ and 0.332 kcal/mol, $R^2=98.1\%$ for the DMS and polarisability model, respectively; 
see also Figure~\ref{f:fig-2}B and \ref{f:fig-2}C).
Nevertheless, the quadratic DMS model yields still very accurate predictions 
of polarization energy and induced multipole moments, even in strong electric field.

Recently, it was found by Fried et~al. by performing molecular dynamics simulations with a classical 
polarisable force field, that the average electric field experienced by water molecule in liquid phase
in room temperature is around 0.014 a.u. ($\sim$70 MV/cm).\cite{Fried.Wang.Boxer.Ren.Pande.JPCB.2013}
This value corresponds to our moderate electric field set of samples (denoted as `M' in Tables~\ref{t:results}
and \ref{t:fields}) for which the
accuracy of the generalised quadratic DMS model is excellent.
Water molecules can experience
stronger electric fields in certain geometries of hydrogen bonded clusters, 
even up to 0.055 a.u. according to study by Reischl et~al.\cite{Reischl.Kofinger.Dellago.MolPhys.2009} 
This electric field magnitude falls into our strong electric field regime (denoted as `S')
in which the generalised quadratic DMS model still provides accurate predictions (see also error distributions 
in Figure~\ref{f:fig-2}B, \ref{f:fig-2}E and \ref{f:fig-2}H). Therefore, it is believed that
the generalised quadratic DMS model is capable of accurately describing 
the polarization of ground\hyp{}state water molecule in inhomogeneous 
electric fields that are comparable in strength to condensed phase conditions.

\section{\label{s:5}Summary and a few concluding remarks}

%In this work, the \emph{ab initio} model of the 
%one\hyp{}particle density matrix susceptibility tensors
%at the HF level was derived for a closed shell molecule 
%in a spatial electric field distribution. It was found 
%that the change in the density matrix element is approximately 
%linearly proportional to the electric field evaluated at the 
%distributed site associated with the centre of charge of the 
%localised molecular orbital. Subsequently, the \emph{ab initio} 
%model served as a basis for the formulation of the generalised 
%model that is, in principle, not limited to the HF theory,
%and takes into account also quadratic and higher\hyp{}order terms 
%with respect to the electric field. The generalised model, which 
%consisted of a set of adjustable parameters, defined the first\hyp{} 
%and second\hyp{}order density matrix susceptibilities that were 
%determined through the least\hyp{}squares linear regression 
%with respect to the training ensemble of electrostatically 
%perturbed states.
%
%The above theoretically developed models were then used 
%to calculate the electric field\hyp{}induced interaction 
%energies, dipole moments and quadrupole moments of water 
%molecule that is subjected to perturbation due to uniform 
%electric field, as well as various point charge environments 
%generating non\hyp{}uniform electric field of weak, moderate
%and strong magnitudes. It was found that, while the \emph{ab initio} model yields qualitatively incorrect
%interaction energies, the atom\hyp{}distributed generalised model can reach quantitative level of accuracy
%for prediction of interaction energies, dipole moments and even quadrupole moments
%once, at least quadratic terms with respect to the electric field, are taken into account. 
In this work, it was demonstrated that the generalised DMS models from Eq.~\eqref{e:final-model.General}
can accurately describe the polarization\hyp{}induced deformation of the one\hyp{}particle ground\hyp{}state electronic density
due to the uniform and non\hyp{}uniform distribution of electric field.
It was also shown that the accuracy of the model can be improved by extending the parameter space of the susceptibility tensors.
Even at the current state of development, the generalised quadratic DMS model
is of comparable and quantitative energetic accuracy as the distributed polarisability model (RMSE of 0.4~kcal/mol in strong 
electric fields and less than 0.06~kcal/mol in moderate and weak fields).
%It is emphasised here that the model can be easily extended by taking into account
%third\hyp{}order terms as well if fine accuracy is required.
%Moreover, the generalized DMS model can be used to any type of one\hyp{}particle density matrices at arbitrary level of theory.
%This however falls out of scope of this contribution and it will be studied
%elsewhere. 
Moreover, any type of one\hyp{}particle density matrix could be treated by using
generalised DMS models, which is much more difficult to realise when a conventional multipole
model is used.
Therefore, it is believed that
the generalised DMS approach
has an important advantage over the multipole\hyp{}based approach, i.e., it provides more detailed
tensor description of the spatial polarisation of the electron density distribution, that could be
combined for example with density\hyp{}based methods of interaction energy calculations.\cite{Mandado.Hermida-Ramon.JCTC.2011}

It is anticipated here that the generalised models from Eq.~\eqref{e:final-model.General}
should be studied in the future with an emphasis 
on their further optimisation and application for molecular aggregates.
For this purpose, more general forms of Eq.~\eqref{e:final-model.General}, as well as the choice of distributed
sites should be further investigated, preferably considering post\hyp{}Hartree\hyp{}Fock methods
and electronically excited states.
Nevertheless, owing to the capability of the proposed approach to correctly capture the 
polarization of ground state's electron density,
the generalised density matrix susceptibility tensors 
could be used in accurate fragment\hyp{}based calculations of extended molecular aggregates
described at correlated levels of theory, particularly including the excited state chemistry.

\begin{acknowledgments}
This project is carried out under POLONEZ programme which has received funding from the European Union's
Horizon~2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement 
No.~665778. \InlineImage{eu-logo.eps} This project is funded by National Science Centre, Poland 
(registration No.~2016/23/P/ST4/01720) within the POLONEZ 3 fellowship. 
%Many thanks to Robert W.~G{\'o}ra and Robert Zale{\'s}ny 
%for reading and commenting on the manuscript.
\end{acknowledgments}

%
\appendix

\section{\label{a:orig-dep} Origin-Independence of the Induced Dipole Moment}

Multipole one\hyp{}electron integrals depend on the origin with respect to which they are evaluated.
However, the induced dipole moment defined in Eq.~\eqref{e:L-projector} 
has to be origin\hyp{}independent. Denoting the overlap AO integrals 
by $S_{\alpha\beta} = \braket{\alpha}{\beta}$,
it follows that $\mathbb{M}_u({\bf r}_Q) = \mathbb{M}_u({\bf 0}) - u_Q {\bf S}$
for $u=x,y,z$. Thus,
%
\begin{equation}
 \mathbb{L}_u({\bf r}_Q) = \mathbb{L}_u({\bf 0})
                          - u_Q {\bf C}^{(0)\dagger} \cdot {\bf S} \cdot \left( {\bf 1} - {\bf D}^{(0)}\right) \;,
\end{equation}
%
where
%
\begin{equation}
 \mathbb{L}_u({\bf 0}) = {\bf C}^{(0)\dagger} \cdot \mathbb{M}_u({\bf 0}) \cdot \left( {\bf 1} - {\bf D}^{(0)}\right) \;.
\end{equation}
%
But
%
\begin{equation}
  {\bf C}^{(0)\dagger} \cdot {\bf S} \cdot \left( {\bf 1} - {\bf D}^{(0)}\right) 
= {\bf C}^{(0)\dagger} \cdot \left( {\bf S} - {\bf 1} \right)
\end{equation}
%
since ${\bf C}^{(0)\dagger} \cdot {\bf S} \cdot {\bf C}^{(0)} = {\bf 1}$. 
In the simplest case of ${\bf S} = {\bf 1}$ (i.e., the AO basis functions are orthogonal) 
$\mathbb{L}_u({\bf r}_Q) = \mathbb{L}_u({\bf 0})$ which guarantees charge conservation.
When a non\hyp{}orthogonal AO basis is used, 
the following general condition must be satisfied:
%
\begin{equation}
 \Re \left\{
  \Tr{
  \left[
  \delta {\bf C} \cdot {\bf C}^{(0)\dagger} \cdot \left( {\bf S} - {\bf 1} \right)  
 \right]}
 \right\} = 0
\end{equation}
%
Due to the fact that ${\BM \Delta}$ considered here is accurate only up to first\hyp{}order, 
the above expression is not valid. 
Therefore, the polarisation\hyp{}induced distributed dipole moments 
defined in Eq.~\eqref{e:L-projector} are exactly origin\hyp{}independent
as long as the AO basis is orthogonalised.
In such a case, one can compute one\hyp{}electron dipole integrals with respect to any origin 
and the resulting susceptibilities $\mathbb{B}^{(i;1)}$ 
from Eq.~\eqref{e:susceptibility-B} will be uniquely defined. They can be back\hyp{}transformed
to the original (non\hyp{}orthogonal) AO basis, 
${\bf X} \cdot \mathbb{B}^{(i;1)} \cdot {\bf X}$, where $\bf X$ is an orthogonaliser.
The numerical tests confirmed that $\Tr\left[\delta{\bf D}\cdot{\bf S}\right] \sim 10^{-16}$
with $\delta{\bf D}$ given by Eq.~\eqref{e:final-model.HF} and susceptibilities transformed
to the non\hyp{}orthogonal AO basis.

\section{\label{a:blocks} Explicit Formulae for Gradient and Hessian Blocks in Linear Regression DMS Model}

The gradient vector ${\bf g}$ and the Hessian matrix ${\bf H}$ 
are built from blocks associated with a particular type of parameters, i.e.,
%
\begin{equation}\label{e:Newton.gH}
 {\bf g} = 
\begin{pmatrix}
{\bf g}^{[1]} \\ 
{\bf g}^{[2]} 
\end{pmatrix} ,\quad
 {\bf H} = 
\begin{pmatrix}
{\bf H}^{[11]} & {\bf H}^{[12]}  \\ 
{\bf H}^{[21]} & {\bf H}^{[22]}  
\end{pmatrix} \;,
\end{equation}
%
where the block indices 1 and 2 correspond to the first\hyp{} and second\hyp{}order susceptibilities, respectively.
Note that the second derivatives of $\delta D^{(N)}$ 
with respect to the adjustable parameters vanish
due to the linear functional form of Eq.~\eqref{e:final-model.General.Parameters}.
Thus, the gradient element of the $r$th block and Hessian element of the $(rs)$th block read
%
\begin{subequations}
 \begin{align}
  g^{[r ]}    &\equiv \frac{\partial   Z}{\partial s^{[r]}} 
     =-2\sum_N \overline{\delta D}^{(N)}
               \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial s^{[r]}} \;,\\
  H^{[rs]} &\equiv \frac{\partial^2 Z}{\partial s^{[r]} \partial s^{[s]}}  
     = 2\sum_N 
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial s^{[r]}}
        \frac{\partial   \left[ \delta D^{(N)} \right]}{\partial s^{[s]}} \;.
 \end{align}
\end{subequations}
%
The explicit formulae for the gradient are
%
\begin{subequations}
 \begin{align}
  g^{[1]}_{ku} &=-2\sum_N \overline{\delta D}^{(N)} F^{(N)}_{ku} \;,\\
  g^{[2]}_{kuw} &=-2r_{uw} \sum_N \overline{\delta D}^{(N)} F^{(N)}_{ku} F^{(N)}_{kw} \;.
 \end{align}
\end{subequations}
%
The Hessian subsequently follows to be
%
\begin{subequations}
 \begin{align}
  H^{[11]}_{ku,lw} &= 2\sum_N F^{(N)}_{ku} F^{(N)}_{lw} \;,\\
  H^{[12]}_{ku,lu'w'} &= 2r_{u'w'} \sum_N F^{(N)}_{ku} F^{(N)}_{lu'} F^{(N)}_{lw'}  \;,\\
  H^{[22]}_{kuw,lu'w'} &= 2r_{uw} r_{u'w'} \sum_N F^{(N)}_{ku} F^{(N)}_{kw} F^{(N)}_{lu'} F^{(N)}_{lw'} \;.
 \end{align}
\end{subequations}
%
Note that due to the symmetry of the Hessian matrix, the block $21$
is a transpose of the block $12$. The composite indices $ku$ and $kuw$ 
are constructed from the distributed site index $k$ and the appropriate 
symmetry\hyp{}adapted ($w<u$) Cartesian component of a particular DMS tensor: 
$u$ for the first\hyp{}order, and $uw$ for the second\hyp{}order susceptibility 
tensor, respectively. The method described above can be easily extended 
to third and higher orders.

% -----------------------
\bibliography{references}
% -----------------------

\end{document}