### 2019-11-29 papers/2_oep_technique.CT/supplementary-r2.tex
SI added to CT paper

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

\title{Supplementary Information\\Ab Initio Effective One-Electron Potential Operators:
Applications for Charge-Transfer Energy in Effective Fragment Potentials}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}

\author{Joanna D. Bednarska}
\author{Marta Cho{\l}uj} 
\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\pacs{}

\maketitle

\tableofcontents

\section{Asymptotic dependence of CT interactions for methanol-water complex}

%
\begin{figure}[h]
\includegraphics[width=0.5\textwidth]{fig-s1.pdf}
\caption{\label{f:fig-s1} {\bf Asymptotic dependence of the charge transfer energy
in the CT/HF0 formulation for water-methanol complex.} 
Water molecule has been translated by $R$ from the starting geometry
along the vector specified in the inset picture.
The counterpoise\hyp{}corrected total interaction energy
is also shown for comparison in light grey color in this figure.
All data were obtained at HF/6-311++G(2df,2pd) level of theory
with the single GDF scheme and aug-cc-pVTZ-jkfit auxiliary basis set.
} 
\end{figure}
%

\section{Statistical validation of OEP method}

% -----------------------
\bibliography{references}
% -----------------------

\end{document}