### 2019-12-30 papers/3_oep_technique.Pauli/supplementary-r3.tex
Figures in Paper 2 modified

\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

% numbering for Figures in SI
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\title{Supplementary Information\\Ab Initio Effective One-Electron Potential Operators. II.
Applications for Exchange-Repulsion Energy in Effective Fragment Potentials}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}

\author{Robert W. G{\'o}ra}
\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\pacs{}

\maketitle

\tableofcontents

\section{Minimalistic auxiliary basis sets for water, methanol and DMSO molecules}

The basis sets are given below. Note that in case of DMSO, the basis set size is smaller than the
minimal basis size (3p orbital on S atoms were not necessary to obtain good fitting in the test basis set).

\begin{verbatim}
[ water ]
cartesian
# primary basis set: 6-311++G** 6D
****
H     0
S   1   1.00
        0.500             1.000000
****
O     0
S   2   1.00
      937.182             0.986162 
        6.892             0.014190 
S   1   1.00
       45.156             1.000000
P   2   1.00
       24.823             0.277579 
        2.783             0.301334 
****

[ methanol ]
cartesian
# primary basis set: 6-311++G** 6D
****
H     0
S   1   1.00
        0.539             1.000000
****
C     0
S   1   1.00
      939.683             1.000000
S   1   1.00
       30.185             1.000000
P   1   1.00
        5.832             1.000000
P   1   1.00
        0.326             1.000000
****
O     0
S   1   1.00
      731.638             1.000000
S   1   1.00
       37.387             1.000000
P   1   1.00
        9.214             1.000000
P   1   1.00
        0.368             1.000000
****

[ dmso ]
cartesian
# primary basis set: 6-31++G** 6D
****
H     0
S   1   1.00
        0.365             1.000000
****
C     0
S   1   1.00
      207.472             1.000000
S   1   1.00
       15.052             1.000000
P   1   1.00
        7.228             1.000000
****
O     0
S   1   1.00
      302.339             1.000000
S   1   1.00
       15.248             1.000000
P   1   1.00
        7.612             1.000000
****
S     0
S   1   1.00
      689.986             1.000000
S   1   1.00
      145.419             1.000000
S   1   1.00
        9.046             1.000000
P   1   1.00
       28.979             1.000000
****
\end{verbatim}

\section{Accuracy of OEP and EFP2 models for exchange-repulsion energy estimation}

%
\begin{figure*}[t]
\includegraphics[width=0.8\textwidth]{fig-2.pdf}
\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of exchange\hyp{}repulsion energy
across various bi\hyp{}molecular systems.} 
(a) NCB31 
database\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
of non\hyp{}covalent interactions
and
(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
was used.
} 
\end{figure*}
%

%
\begin{figure*}[t]
\includegraphics[width=0.8\textwidth]{fig-s2.pdf}
\caption{\label{f:fig-2} {\bf Accuracy of the OEP and EFP2 models of first\hyp{}order repulsion energy
across various bi\hyp{}molecular systems.} 
(a) NCB31 
database\cite{Zhao.Schultz.Truhlar.JCTC.2006,
Zhao.Truhlar.JCTC.2005,Zhao.Schultz.Truhlar.JCTC.2006,Zhao.Schultz.Truhlar.JCP.2005} 
of non\hyp{}covalent interactions
and
(b) BBI subset\cite{Burns.Faver.Zheng.Marshall.Smith.Vanommeslaeghe.MacKerell.Merz.Sherrill.JCP.2017} 
of backbone\hyp{}backbone interactions in proteins from the BioFragment Database.
For the OEP calculations, the EDF-1 scheme with the aug-cc-pVDZ-jkfit auxiliary basis set
was used.
} 
\end{figure*}
%

\clearpage
\section{References}

% -----------------------
\bibliography{references}
% -----------------------

\end{document}