### 2020-02-21 papers/2_oep_technique.CT/supplementary-r2.tex

\documentclass[aip,jcp,amsmath,amssymb,preprint,floatfix]{revtex4-1}
\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

% numbering for Figures in SI
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\title{Supplementary Information\\Ab Initio Effective One-Electron Potential Operators:
Applications for Charge-Transfer Energy in Effective Fragment Potentials}

\author{Bartosz B{\l}asiak}
\email[]{blasiak.bartosz@gmail.com}
\homepage[]{https://www.polonez.pwr.edu.pl}

\author{Joanna D. Bednarska}
\author{Marta Cho{\l}uj} 
\author{Wojciech Bartkowiak}

\affiliation{Department of Physical and Quantum Chemistry, Faculty of Chemistry, 
Wroc{\l}aw University of Science and Technology, 
Wybrze{\.z}e Wyspia{\'n}skiego 27, Wroc{\l}aw 50-370, Poland}

\date{\today}

\pacs{}

\maketitle

\tableofcontents

\section{Asymptotic dependence of CT interactions for methanol-water complex}

%
\begin{figure}[h]
\includegraphics[width=0.5\textwidth]{fig-s1.pdf}
\caption{\label{f:fig-s1} {\bf Asymptotic dependence of the charge transfer energy
in the CT/HF0 formulation for water-methanol complex.} 
Water molecule has been translated by $R$ from the starting geometry
along the vector specified in the inset picture.
The counterpoise\hyp{}corrected total interaction energy
is also shown for comparison in light grey color in this figure.
All data were obtained at HF/6-311++G(d,p) level of theory
with the EDF-1 scheme and aug-cc-pVQZ-jkfit auxiliary basis set.
} 
\end{figure}
%

\section{Statistical validation of OEP method}

In Figure~\ref{f:fig-s2} interaction energy components
obtained by using the code developed in this work are
compared to the ones obtained by using the EFP2 calculation routines
in GAMESS US quantum chemistry package\cite{GAMESS.JCC.1993} (version: Sept 30, 2017 R2 Public Release). 
For this, EFP2 parameters
were generated by running the \texttt{MAKEFP} routine
with \texttt{CTVVO} option switched to \texttt{.FALSE.} which corresponds
to using all the canonical molecular orbitals from HF calculations
instead of valence virtual orbitals.
Generally, Coulombic energies (Figure~S2(a)) 
are in very good agreement with a few exeptions in the PPS5 set,
for which the differences are due to the inaccuracies in the DMTP
expansion used in the EFP2 model -- note that the Coulombic component in the DDS/HF method
is free from multipole approximation and is equivalent to the first\hyp{}order SAPT Coulombic energy.
Exchange\hyp{}repulsion and induction energies (Figures~S2(b) and S2(c), respectively)
are in very good agreement. CT energies (Figure~S2(d)) 
evaluated with CAMM up to quadrupoles for the potential energy matrix elements (black filled circles)
are generally also in decent quantitative agreement,
except for NH$_3$--FCl, (HCOOH)$_2$ and (HCONH$_2$)$_2$ 
%and parallel\hyp{}displaced benzene dimer
where the differences
are between 2--4 kcal/mol. Only in the case of the latter three systems,
adding distributed octupoles (blue crosses in Figure~S2(d)) 
lowers the CT energy values by around 1 kcal/mol,
with negligible changes for the rest.

%
\begin{figure}[h]
\includegraphics[width=0.8\textwidth]{fig-s2.pdf}
\caption{\label{f:fig-s2} {\bf Validation of implementation of the interaction energy components
with respect to the EFP2 GAMESS US code.} 
Results obtained for the NCB31 database set assuming 6-311++G(d,p) primary basis set. 
(a) - Coulombic, (b) - exchange-repulsion, (c) - induction, and (d) - charge-transfer
interaction energy. In (d), black circles and blue crosses correspond to the
CT EFP2 energies with CAMM up to distributed quadrupoles and octupoles, respectively.
For more details see the main text (Section III, \emph{Calculation Details}).
} 
\end{figure}
%

% -----------------------
\bibliography{references}
% -----------------------

\end{document}